Article Title

### 2014-12-17 final.tex
final revision in final.tex

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf sample}$] -- the sample covariance matrix, the unbiased estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \emph{i.e.}\;toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

@@

We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for representing the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. In our data, the sample correlation coefficients were largely positive and low.  We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.  This estimator revealed a sparse network of partial correlations (`interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons. We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, the average partial correlations and the inferred positive interactions in particular exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations. In contrast, the inferred negative interactions were less selective.

% Results and Discussion can be combined.
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\section*{Results}
\subsection*{Covariance estimation}
The covariance matrix is defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
    \end{equation}
    where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons in a time bin of some duration; and $\mu$ is the vector of expected firing rates.

Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
\begin{equation}\label{eq:sample}
    C_{\sf sample} = \frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
    \end{equation}
where $n$ is the number of time bins in $T$. 
The sample covariance matrix yields an unbiased estimate of the true covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean activity, $\bar x$, is estimated independently, in a separate data sample. 
In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the means are estimated independently.

@@

The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}. Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of pairwise interactions \cite{Whittaker:1990}. For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}. More generally, partial correlations can serve as a measure of conditional independence under the assumption that dependencies in the system are close to linear effects \cite{Whittaker:1990,Baba:2004}. As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.

@@

Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix of a low rank and $D$ is a diagonal matrix of independent variances. This approximation is the basis for \emph{factor analysis} \cite{Anderson:2003}. The estimator,
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = L + (1-\lambda)D + \lambda \bar D,
\end{equation}
has two hyperparameters: the rank of $L$ (\emph{i.e.}\;the number of latent factors) and the shrinkage intensity $\lambda$ to shrink the independent variances toward their mean value $\bar D$. The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (Fig.~\ref{fig:1} Row 1, B).

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$:
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \emph{i.e.}\;one in which a large fraction of off-diagonal elements are set to zero.  
The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$. 
The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. 
The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1} Row 1, C).

Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component:
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1},
\end{equation}
where $S$ is a sparse matrix and $L$ is a low-rank matrix. The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$. See Methods for a more detailed explanation. The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (Fig.~\ref{fig:1} Row 1, D) \cite{Chandrasekaran:2010,Ma:2013}.

@@

Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, the validation loss is an unbiased estimate of the true loss, up to a constant:
\begin{equation}
    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
\end{equation}
Therefore, validation procedure allow comparing the relative values of the average loss attained by different covariance matrix estimators.

We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
The validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators even without access to the ground truth (Fig.~\ref{fig:1} Row 6).

@@

In Gaussian models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data. 
To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
Conveniently, the Ising model has equivalent mathematical form to the Gaussian distribution,
\begin{equation}\label{eq:ising}
    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right)
\end{equation}
but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives, although they do not have a simple relationship to the means and the covariance matrix. 
Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.

@@

This simulation study demonstrated that cross-validated evaluation of regularized estimators of the covariance matrix of population activity can discriminate between structures of dependencies in the population. The identity of the most efficient covariance estimator for a particular neural circuit is therefore an empirical finding characterizing the nature of circuit  interactions.

\subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
We recorded the calcium activity of densely sampled populations of neurons in layers 2/3 and upper layer 4 in primary visual cortex of sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}. This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:3} C and D).  The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:3} C). Only cells that produced detectable calcium activity were included in the analysis (see Methods).  First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.  Each grating was displayed for 500 ms, without intervening blanks.  This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:3} D). To estimate the noise correlation matrix, we presented only two distinct directions in some experiments or five directions in others with 100--300 repetitions of each direction. Each grating lasted 1 second and was followed by a 1-second blank.  The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the correlation matrix.   The sample correlation coefficients were largely positive and low (Fig.~\ref{fig:3} E and F). The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (Fig.~\ref{fig:6} D).

@@

To test this hypothesis, we computed the relative validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}). The respective median differences of the validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

\subsection*{Structure of $C_{\sf sparse+latent}$ estimates}
We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:5} and Fig.~\ref{fig:6}). Although the regularized estimates were similar to the sample correlation matrix (Fig.~\ref{fig:5} A and B), the corresponding partial correlation matrices differed substantially (Fig.~\ref{fig:5} C and D). The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (Fig.~\ref{fig:5} E). The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (Fig.~\ref{fig:5} F, G, H, and I).

@@

Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013, Sadovsky:2014}.  We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to the same level of connectivity. The networks revealed by the two methods differed substantially. In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (Fig.~\ref{fig:5} F, H, and I). In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:5} F) where high correlations should be expected given the rest of the correlation matrix.

@@

We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:7}).  Five sites with highest pairwise connectivities were included in the analysis. Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component. Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.

First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta \mbox{ori}$) of pairs of significantly ($\alpha=0.05$) tuned cells. The partial correlations decayed more rapidly with $\Delta\mbox{ori}$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients). Positive connectivity decreased with $\Delta\mbox{ori}$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:7} D): The slope in the logistic model of connectivity with respect to $\Delta\mbox{ori}$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).

Second, we compared how correlations and connectivity depended on the physical distance separating pairs of cells. We distinguished between lateral distance, $\Delta x$, in the plane parallel to the pia, and vertical distance, $\Delta z$, orthogonal to the pia.  When considering the dependence on $\Delta x$, the analysis was limited to cell pairs located at the same depth with $\Delta z < 30\,\mu\mbox{m}$; conversely, when considering the dependence on $\Delta z$, only vertically aligned cell pairs with $\Delta x < 30\,\mu\mbox{m}$ were included. Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (Fig.~\ref{fig:7} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$. Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data. The negative connectivity decayed slower with distance (Fig.~\ref{fig:7} E and F): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).

\section*{Discussion}
\subsection*{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often represented as a graph of pairwise interactions. The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input. Therefore, conditional independence can signify the absence of a direct causal influence.

@@

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013, Sadovsky:2014}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt them here.

@@

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013, Theis:2014}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than electrophysiological recordings.

@@

Let $T_c$ and $T_c^\prime$ denote the time bin indices for the training and testing samples, respectively, limited to condition $c$. Here, the prime symbol marks quantities estimated from the testing sample.

>>>>

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf sample}$] -- the sample covariance matrix, the unbiased estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \emph{i.e.}\;toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;all but a small fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

@@

We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for representing the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. In our data, the sample correlation coefficients were largely positive and low.  We found that the most efficient estimate of the correlation matrix in these data was $C_{\sf sparse+latent}$.  This estimator revealed a sparse network of partial correlations (`interactions'), between the observed neurons; it also inferred a number of latent units interacting with the observed neurons. We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, the inferred positive interactions exhibited a substantially stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations. In contrast, the inferred negative interactions were less selective.

% Results and Discussion can be combined.
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\section*{Results}
\subsection*{Covariance estimation}
The covariance matrix is defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
    \end{equation}
    where the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons in a time bin of some duration, $\E{\cdot}$ denotes expectation, and $\mu$ is the vector of expected firing rates.

Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ contains observed firing rates in time bin $t$, and an independent estimate of the mean firing rates $\bar x$, the \emph{sample covariance matrix},
\begin{equation}\label{eq:sample}
    C_{\sf sample} = \frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
    \end{equation}
where $n$ is the number of time bins in $T$, is an unbiased estimate of the true covariance matrix, \emph{i.e.} $\E{C_{\sf sample}}=\Sigma$. 
In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the mean activity is estimated independently from a separate sample.

@@

The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}. Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of pairwise interactions \cite{Whittaker:1990}. For multivariate Gaussian distributions, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}. More generally, partial correlations can serve as a measure of conditional independence under the assumption that dependencies in the system are close to linear effects \cite{Whittaker:1990,Baba:2004}. As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.

@@

Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model, The estimator,
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = L + D,
\end{equation}
where $L$ is a $p\times p$ symmetric positive semidefinite matrix with low rank and $D$ is a diagonal matrix. 
This approximation is the basis for \emph{factor analysis} \cite{Anderson:2003}, where matrix $L$ represents covariances arising from latent factors. The rank of $L$ corresponds to the number of latent factors. Matrix $D$ contains the variances of the cells' independent activity from the latent factors. The estimator is regularized by selecting the rank of $L$ and by shrinking the independent variances in $D$ toward their mean.
The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (Fig.~\ref{fig:1} Row 1, B).

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$:
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \emph{i.e.}\;one in which a large fraction of off-diagonal elements are set to zero.  
The estimator is regularized by adjusting the sparsity (fraction of off-diagonal zeros) of $S$. 
The problem of finding the optimal set of non-zero elements in $S$ is known as \emph{covariance selection} \cite{Dempster:1972}. 
The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1} Row 1, C).

Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the difference of a sparse component and a low-rank component:
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1},
\end{equation}
where $S$ is a sparse matrix and $L$ is a low-rank matrix. The estimator is regularized by adjusting the sparsity of $S$ and the rank of $L$. See Methods for more detailed explanations. The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (Fig.~\ref{fig:1} Row 1, D) \cite{Chandrasekaran:2010,Ma:2013}.

@@

Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, the validation loss is an unbiased estimate of the true loss, up to a constant:
\begin{equation}
    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
\end{equation}
Therefore, validation procedures allow comparing the relative values of the loss attained by different covariance matrix estimators.

We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Rows 4 and 5).
The validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the estimators even without access to the ground truth (Fig.~\ref{fig:1} Row 6).

@@

In Gaussian models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data. 
To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
Conveniently, the Ising model has equivalent mathematical form to the Gaussian distribution,
\begin{equation}\label{eq:ising}
    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right)
\end{equation}
but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
The partition function $Z(J,h)$ normalizes the distributions on the models' respective domains. 
In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives, although they do not have a simple relationship to the means and the covariance matrix. 
Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.

@@

This simulation study demonstrated that cross-validated evaluation of regularized estimators of the covariance matrices of population activity can discriminate between structures of dependencies in the population. The selection of the most efficient covariance estimators for particular neural circuits is therefore an empirical finding characteristic of the nature of circuit interactions.

\subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
We recorded the calcium activity of densely sampled populations of neurons in layers 2/3 and upper layer 4 in primary visual cortex of sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}. This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in $200\times200\times100$ $\mu$m$^3$ volumes of cortical tissue (Fig.~\ref{fig:3} C and D).  The instantaneous firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:3} C). Only cells that produced detectable calcium activity were included in the analysis (see Methods).  First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.  Each grating was played for 500 ms, without intervening blanks.  This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:3} D). To estimate the noise correlation matrix, we presented only two distinct directions in some experiments or five directions in others with 100--300 repetitions of each direction. Each grating lasted 1 second and was followed by a 1-second blank.  The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the correlation matrix.   The sample correlation coefficients were largely positive and low (Fig.~\ref{fig:3} E and F). The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018.

@@

To test this hypothesis, we computed the validation loss of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}). The respective median differences of the validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, Wilcoxon signed rank test).

\subsection*{Structure of $C_{\sf sparse+latent}$ estimates}
We examined the composition of the $C_{\sf sparse+latent}$ estimates for each imaged site (Fig.~\ref{fig:5} and Fig.~\ref{fig:6}). Although the regularized estimates were similar to the sample correlation matrix (Fig.~\ref{fig:5} A and B), the corresponding partial correlation matrices differed substantially (Fig.~\ref{fig:5} C and D). The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (Fig.~\ref{fig:5} E). The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (Fig.~\ref{fig:5} F, G, H, and I).

@@

Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013, Sadovsky:2014}.  We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to the same level of connectivity. The networks revealed by the two methods differed substantially. In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (Fig.~\ref{fig:5} F, H, and I). In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:5} F) where high correlations are expected given the rest of the correlation matrix.

@@

We then examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:7}).  Five sites with highest pairwise connectivities were included in the analysis. Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component. Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, segregated into positive and negative connectivities.

First, we analyzed how correlations and connectivity depended on the differences in preferred orientations ($\Delta \mbox{ori}$) of pairs of significantly ($\alpha=0.05$) tuned cells. The partial correlations decayed more rapidly with $\Delta\mbox{ori}$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients in normalized data). Positive connectivity decreased with $\Delta\mbox{ori}$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:7} D): The slope in the logistic model of connectivity with respect to $\Delta\mbox{ori}$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).

Second, we compared how correlations and connectivity depended on the physical distance separating pairs of cells. We distinguished between lateral distance, $\Delta x$, in the plane parallel to the pia, and vertical distance, $\Delta z$, orthogonal to the pia.  When considering the dependence on $\Delta x$, the analysis was limited to cell pairs located at the same depth with $\Delta z < 30\,\mu\mbox{m}$; conversely, when considering the dependence on $\Delta z$, only vertically aligned cell pairs with $\Delta x < 30\,\mu\mbox{m}$ were included. Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient in normalized data) (Fig.~\ref{fig:7} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$. Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data. The negative connectivity decayed slower with distance (Fig.~\ref{fig:7} E and F): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).

\section*{Discussion}
\subsection*{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often represented as a graph of pairwise interactions. The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input. Therefore, conditional independence shown statistically can signify the absence of a direct causal influence.

@@

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013, Sadovsky:2014}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt applying graph-theoretical analyses to our results.

@@

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by the calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013, Theis:2014}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than electrophysiological recordings.

@@

Let $T_c$ and $T_c^\prime$ denote the sets of time bin indices for the training and testing samples, respectively, limited to condition $c$.
